---
title: "README - Wrangling Cannabis Data Sets"
author: "Martin Frigaard"
date: "current version: `r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
require(tidyverse)
library(plotly)
require(janitor)
require(skimr)
library(mosaic)
library(inspectdf)
library(visdat)
library(DT)
library(hrbrthemes)
# base options ----
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 999999
)
# knitr chunk options ----
knitr::opts_chunk$set(
  echo = FALSE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  fig.width = 9, # figure width
  fig.height = 6, # figure height
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/"
) # location of files
# knitr knit settings ----
knitr::opts_knit$set(
  width = 78
)
# set theme
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc(
  base_size = 11,
  strip_text_size = 12,
  axis_title_size = 14,
  plot_title_size = 21,
  subtitle_size = 18
))
```

# Motivation

This script outlines the steps to wrangle the Cannabis data. We initially used an open online retail data set from the [UCI Machine Learning Laboratory](https://archive.ics.uci.edu/ml/datasets/online+retail) for the week-over-week visualizations (`OnlineRetail`), and combined this with categories from the [`kushyapp`](https://github.com/kushyapp/cannabis-dataset).

# The data 

All of the data are in the `data/` folder. 

```{r data-folder, message=FALSE, warning=FALSE}
fs::dir_tree("data/", recurse = FALSE)
```

## Import the Kushy data 

These data sets provide the categories and brands from the Cannabis industry. 

```{r kushy_files}
kushy_files <- fs::dir_ls("data/kushy-datasets/csv", regexp = "kushy_")
kushy_files
```

Import these files.

```{r import-kushy-data, message=FALSE, warning=FALSE}
KushyBrands <- readr::read_csv(kushy_files[1])
KushyProducts <- readr::read_csv(kushy_files[2])
KushyShops <- readr::read_csv(kushy_files[3])
KushyStrains <- readr::read_csv(kushy_files[4])
```

## Kushy Brands

These are the brands in the `KushyBrands` data frame. I want to see how many unique brands are represented here, but first I need to know what I am dealing with in terms of completeness. I can check this with `skimr`

```{r KushyBrands-vis-miss}
KushyBrands %>% 
  visdat::vis_miss() + 
  coord_flip()
```


This shows I can remove the `instagram` variable, and the `category` variable is missing `406` values. Hopefully this won't an issue. We can do without the `slug` too. 

```{r remove-slug-instagram}
KushyBrands <- KushyBrands %>% 
  # remove the instagram and slug variables
  dplyr::select(-c(instagram, slug)) %>% 
  dplyr::rename(mult_category = category,
                brand_name = name)
KushyBrands %>% 
  dplyr::select_if(is.character) %>% 
  inspectdf::inspect_cat() %>% 
  inspectdf::show_plot()
```

Now I can check the unique `name`s for each brand. These are stored in the `brand_name` variable.

### The Kushy brand names

These are the brand names in the `KushyBrands` data frame.

```{r name-unique-length}
KushyBrands %>% 
  dplyr::count(brand_name, sort = TRUE) 
```

The output above shows us there are `1776` unique observations in the `name` variable, but two of them (`Brewbudz by Cannabiniers` and `Dixie Elixirs & Edibles`) appear more than once. 

### Kushy brand categories 

When we look at the categories we see that there are multiple categories (`mult_category`) crammed into a single column separated by commas. 

```{r head-KushyBrands}
KushyBrands %>% 
  utils::head(10)
```

We want to separate each category out into an individual rows, because `category` *might* be useful for itemizing products later. The categories are currently separated by commas, but we can split them up and separate the rows with `tidyr::separate_rows()`. We also need to remove the whitespace from the `category` variable, and remove any rows (with other quotes). 

```{r create-TidyBrandCat}
TidyBrandCat <- KushyBrands %>% 
  # split across rows
  tidyr::separate_rows(data = ., mult_category, sep = ",") %>%
  # trim the whitespace from mult_category to category
  dplyr::mutate(category = stringr::str_trim(mult_category)) %>%
  # remove the empty rows
  dplyr::filter(category != "") %>% 
  # remove the mult_category
  dplyr::select(-mult_category,
                brand_category = category)
TidyBrandCat %>% dplyr::glimpse(78)
```

We've just changed the number of rows in this data frame from `1778` to `1988` because we've added a new row for each category, and these were previously on a single row (separated by a commas). If we count the number of names now, we can see there are more counts in the `n` column.

```{r recount-names}
TidyBrandCat %>% 
  dplyr::count(brand_name, sort = TRUE) %>% 
  utils::head(10)
KushyBrands %>% 
  dplyr::count(brand_name, sort = TRUE) %>% 
  utils::head(10)
```

These are the counts for each name (products and brands), and we can see there are fewer names in the `TidyBrandCat` data frame than the `KushyBrands` data frame. 

#### Verify the BaM case study

Check the `BaM` case study to see if there are multiple categories for this brand. 

```{r wide-BaM}
KushyBrands %>% 
  dplyr::filter(brand_name == "BaM")
```

And we can see there are four comma-separated categories in the `KushyBrands` data, and the same four categories are in the `TidyBrandCat`. 

```{r tidy-BaM}
TidyBrandCat %>% 
  dplyr::filter(brand_name == "BaM")
```

This shows us there are five categories for `BaM` in `KushyBrands`, and five categories for `BaM` in `TidyBrandCat`. Now we can check the top brand categories.

## Top Brand Categories

When we check the top categories in the brands table, we see there are mostly `Concentrates`, `Flower`, and `Edibles`. 

```{r glimpse-TidyBrands}
TidyBrandCat %>% 
  dplyr::count(brand_category, sort = TRUE) %>% 
  dplyr::rename(category_count = n) %>% 
  utils::head(10) %>% 
  ggplot2::ggplot(aes(x = category_count, 
                      y = forcats::fct_reorder(.f = factor(brand_category), 
                                               .x = category_count))) + 
  ggplot2::geom_col(aes(fill = brand_category), 
           show.legend = FALSE) + 
  labs(title = "Top 10 Kushy brand categories" , 
       y = "Brand Category", x = "Count")
```

We will still want to use the `location` to replace the `country` variable in the online retail data.

## Unique locations

These are the unique locations in the `TidyBrandCat` dataset. 

```{r count-location}
library(magrittr)
TidyBrandCat %$% base::unique(location)
```

We will make this a little cleaner for the graphs. 

```{r format-location}
TidyBrandCat <- TidyBrandCat %>% 
  dplyr::mutate(location = case_when(
    # alberta 
    location %in% c("AB", "ab", "Ab") ~ "Alberta",
    # manitoba
    location %in% c("MB", "mb", "Mb") ~ "Manitoba",
    # british columbia 
    location %in% c("BC", "British Columbia") ~ "British Columbia",
    # pleasant hill assuming this is Saskatchewan
    location %in% c("Pleasant Hill") ~ "Saskatchewan",
    # nova scotia
    location %in% c("NS") ~ "Nova Scotia",
    # ontario 
    location %in% c("ON") ~ "Ontario",
    # newfoundland
    location %in% c("NL") ~ "Newfoundland",
    # california 
    location %in% c("CA", "Ca", "ca", "CALIFORNIA") ~ "California",
    # colorado
    location %in% c("CO", "Colorado", "Co") ~ "Colorado",
    # washington
    location %in% c("WA", "Washington", "Wa") ~ "Washington",
     # arizona
    location %in% c("AZ", "Arizona", "Az") ~ "Arizona",
    # nevada
    location %in% c("NV", "Nv") ~ "Nevada",
    # michigan
    location %in% c("MI", "Michigan") ~ "Michigan",
    # oregon
    location %in% c("OR", "Oregon", "Or") ~ "Oregon",
    # Sp 
    location %in% c("Sp", "sp") ~ "Salisbury Postcode Area (UK)",
    # all else
    TRUE ~ location
  )) 
```

Here we can visualize the top brand categories by their `location`.

```{r viz-top-brand-categories}
TidyBrandCat %>% 
  count(brand_category, sort = TRUE) %>% 
  rename(category_count = n) %>% 
  dplyr::inner_join(x = ., 
                    y = TidyBrandCat, 
                    by = "brand_category") %>% 
  ggplot2::ggplot(aes(x = category_count, 
                      y = forcats::fct_reorder(.f = factor(brand_category), 
                                               .x = category_count),
                      group = location)) + 
  ggplot2::geom_col(aes(fill = brand_category), show.legend = FALSE) + 
  ggplot2::facet_wrap(. ~ location, 
             scales = "free_x") + 
  ggplot2::labs(title = "Top Kushy brand categories by location" , 
       y = "Brand Category", x = "Count")
```

Now we can see the vast majority of these are from `California`. We will look to join this with the product data below. We'll reduce this dataset to only the 

```{r TidyBrandCat}
TidyBrandCat <- TidyBrandCat %>% 
  dplyr::select(-id)
```


***

## Kushy Product Data 

The kushy products are stored in the `KushyProducts` data frame. These might be a better candidate for binding to the online retail data. 

```{r KushyProducts}
KushyProducts %>% 
  visdat::vis_miss() + 
  coord_flip()
```

We can remove the `slug`, `strain`, `thc`, `cbd`, `lab_test`.

```{r remove-missing-vars}
KushyProducts <- KushyProducts %>% 
  dplyr::select(-slug, 
                -strain, 
                -thc, 
                -cbd,
                -lab_test) 
KushyProducts %>% glimpse()
```

How we can check the character variables in the data frame. 

```{r KushyProducts-inspect_cat}
KushyProducts %>% 
  dplyr::select_if(is.character) %>% 
  inspectdf::inspect_cat() %>% 
  inspectdf::show_plot(text_labels = TRUE)
```

We will check the count of `name`s first. 

#### Product `name`

The `name` for each product is below. 

```{r name-product}
KushyProducts %>% 
  dplyr::count(name, sort = TRUE) %>% 
  utils::head(10)
```

### Separate product name and product details

I am going to separate these terms by the hyphen ` -` and create a `product_name` variable and `product_details` variable. These two can be used to clean up the products into categories and their names.

```{r create-new-product_name-and-product_details}
KushyProducts <- KushyProducts %>% 
  tidyr::separate(col = name, 
                  into = c("product_name", "product_details"), 
                  sep = " -",
                  remove = FALSE) %>% 
  dplyr::mutate(product_details = stringr::str_trim(product_details))
# remove the empty variables
KushyProducts <- KushyProducts %>% 
  dplyr::select(id, 
                name, 
                product_name, 
                product_details, 
                brand,
                product_category = category) 
KushyProducts %>% 
  dplyr::select(name,
                dplyr::contains("product")) %>% 
  utils::head(10)
```

We now have two variables, one for `product_name` and one for `product_category`. These each contain the same information, but arranged in different ways. 

#### Products: `category`

These are the categories above.

```{r product-categories}
KushyProducts %>% 
  dplyr::count(product_category, sort = TRUE)
```

### Separate product categories across rows 

We need to clean the categories just like we did in the brands data frame. 

```{r clean-categories}
TidyProdCat <- KushyProducts %>% 
  # split across rows
  tidyr::separate_rows(data = ., product_category, sep = ",") %>%
  # trim the whitespace from mult_category to category
  dplyr::mutate(category = stringr::str_trim(product_category)) %>%
  # remove the empty rows
  dplyr::filter(product_category != "") %>% 
  # reorganize
  dplyr::select(id, 
                name,
                product_name, 
                product_details, 
                product_category,
                dplyr::everything()) 
TidyProdCat %>% 
  utils::head(10)
```

We can check these the same way we did with the brand categories, by selecting a case. We will use the `"Casey Jones - Crumble"` case to see if the `category` has been separated. 

The original `KushyProducts` dataset:

```{r casey-jones}
KushyProducts %>% 
  dplyr::filter(name == "Casey Jones - Crumble") %>% 
  dplyr::select(name, product_category)
```
This should be translated into three rows in the `TidyProdCat` dataset. 

```{r tidy-casey-jones}
TidyProdCat %>% 
  dplyr::filter(name == "Casey Jones - Crumble") %>% 
  dplyr::select(name, product_category)
```

And it is!

#### Products: `id`

This looks like a unique `id` for each product in the data frame.

```{r check-ids}
identical(x = KushyProducts %>% dplyr::distinct(id) %>% base::nrow(),
          y = KushyProducts %>% base::nrow())
```

So each `id` is unique, and there is one per row. We know this is not the case in the new `TidyProdCat` data frame. 

#### Products: `brand`

This appears similar to the `brand` variable, but it lists many more levels. 

```{r total-distinct-names}
TidyProdCat %>% dplyr::distinct(brand) %>% nrow()
```

These look similar to the brands in the `brand_name` from `TidyBrandCat`.

```{r TidyBrandCat-unique-brand_names}
TidyBrandCat %$%
  base::unique(brand_name) %>% 
  head(50)
```

```{r unique-brand-TidyProdCat}
TidyProdCat %$% 
  base::unique(brand) %>% 
  head(50)
```

### Remove the `Unknown Producer`

Let's filter out the `"Unknown Producer"`s because they will be of no use.  

```{r brand-product}
TidyProdCat <- TidyProdCat %>% 
  dplyr::filter(brand != "Unknown Producer") 
TidyProdCat %>% dplyr::glimpse(78)
```

This shows there are `25,862` rows in the `TidyProdCat` dataset. 

```{r count-product_name}
TidyProdCat %>% 
  dplyr::count(product_name, sort = TRUE)
```

This shows there are some product names that occur much more than others, so we will assign these counts to the data `TidyProdCat`frame.  

## Create `product_name` counts variable

First I will get a data frame with only the distinct names of each product, sorted with the most common products first. 

```{r TidyProdCount}
TidyProdCount <- TidyProdCat %>%
  # count the product (name)
  dplyr::count(product_name, sort = TRUE) %>% 
  # join back to the original table
  dplyr::inner_join(x = ., 
                    y = TidyProdCat, 
                    by = "product_name") %>% 
  # rename n column to count
  dplyr::rename(product_name_count = n) %>% 
  # remove the strain, name, thc, and id
  dplyr::select(contains("product"),
                brand_name = brand) %>% 
  # arrange by brand count
  dplyr::arrange(desc(product_name_count)) %>% 
  # get the distinct rows 
  dplyr::distinct(.keep_all = TRUE)
# check the dimensions
TidyProdCount %>% dplyr::glimpse(78)
```

We will check the number of `brands` with `base::length(base::unique())`.

```{r check-brand_name}
length(unique(TidyProdCount$brand_name))
length(unique(TidyBrandCat$brand_name))
```

These numbers are close, so hopefully I will get a descent number when I join this to the brands. 

Now I can join this to the `TidyProdCount` dataset on `brand_name`

### Join products and brands

This joins the brands and products on `brand_name`.

```{r join-TidyProdCount-TidyBrandCat}
TidyProdBrands <- TidyProdCount %>% 
  dplyr::inner_join(x = ., 
                    y = TidyBrandCat, 
                    by = c("brand_name")) %>% 
  dplyr::select(contains("product"),
                contains("brand"),
                location)
```

### Create product ranks

The `product_name_count` variable is a way to rank the products, so I will use the `base::rank()` function to create a `product_rank` variable, too.

```{r TidyProdBrandRank}
TidyProdBrandRank <- TidyProdBrands %>% 
  dplyr::distinct(product_name, product_name_count) %>% 
  dplyr::mutate(product_rank =  base::order(product_name_count, 
                                           na.last = TRUE, 
                                           decreasing = TRUE)) %>% 
  dplyr::select(product_name, product_name_count, product_rank)
utils::head(TidyProdBrandRank)
utils::tail(TidyProdBrandRank)
```

Now I can join this new `product_rank` back to the `TidyProdCount` dataset 

```{r TidyProdRank}
TidyProdRank <- TidyProdBrands %>% 
  dplyr::left_join(x = ., 
                   y = TidyProdBrandRank, 
                   by = c("product_name", "product_name_count")) %>% 
  # select only useful vars here
  dplyr::select(product_name,
                product_rank, 
                product_category,
                product_details,
                brand_name,
                location) %>% 
  # get distint rows
  dplyr::distinct()
TidyProdRank %>% glimpse(78)
```

Now we have a distinct row for each product, the product ranking, the `product_category` category, and the product details.

### Generate a set of random dates 

We can create a `sim_datetime()` function to create a series of dates. 

```{r random_datetime}
sim_datetime <- function(n, start_time = "2019/01/01", end_time = "2020/01/01") {
  # convert to date (as.POSIXct)
  start_time <- base::as.POSIXct(as.Date(start_time))
  end_time <- as.POSIXct(as.Date(end_time))
  # get datetime difference 
  time_diff <- base::as.numeric(base::difftime(end_time, start_time, unit = "sec"))
  # get random (runif) intervals
  runif_time_interval <- base::sort(stats::runif(n = n, min = 0, max = time_diff))
  # return_time is start_time and runif_time_interval
  return_time <- start_time + runif_time_interval
  
  return(return_time)
  
}
```

Now create these and assign to `TidyProdRank`

```{r assign-invoice_dates}
# create random dates
invoice_dates <- sim_datetime(n = base::nrow(TidyProdRank), 
                              start_time = "2018-01-01", 
                              end_time = "2019-12-31")
# assign to TidyProdRank
TidyProdRank <- invoice_dates %>% 
  tibble::as_tibble(x = .,) %>% 
  dplyr::rename(invoice_date = value) %>% 
  dplyr::bind_cols(TidyProdRank)
# remove any before 2018
TidyProdRank <- TidyProdRank %>% 
  dplyr::filter(invoice_date >= lubridate::as_date("2018-01-01"))
```

### Generate random sales data for `product_name` per `product_details`

Now we want a dollar amount for the items sold (`product_price`), which is a combination of `product_name` and `product_details`.

```{r unique-product_name-product_details}
length(unique(TidyProdRank$product_name))
length(unique(TidyProdRank$product_details))
```

This will get the distinct combinations of product names and details, then generate a variable of `product_price`s, and join this back to `TidyProdRank`.

```{r product_price}
ProdPriceDetails <- TidyProdRank %>% 
  # get names and details
  dplyr::select(product_name, product_details) %>% 
  # get distinct rows
  dplyr::distinct() 
# ProdPriceDetails %>% base::nrow()
product_price <-  runif(n = base::nrow(ProdPriceDetails), 
                        min = 4.99, 
                        max = 99.99)
ProdPrices <- product_price %>% 
  tibble::as_tibble() %>% 
  dplyr::rename(product_price = value) %>% 
  dplyr::bind_cols(ProdPriceDetails)
TidyProdRank <- TidyProdRank %>% 
  dplyr::left_join(x = ., 
                   y = ProdPrices, 
                   by = c("product_name", "product_details")) %>% 
  dplyr::mutate(product_price = base::round(product_price, digits = 2))
```

### Generate a number of `units_sold`

We will create a variable of `units_sold`, one for each row in `TidyProdRank`

```{r UnitsSold}
units_sold <- runif(n = base::nrow(TidyProdRank), 
                        min = 1L, 
                        max = 10L)
UnitsSold <- units_sold %>% 
  as_tibble() %>% 
  rename(units_sold = value) %>% 
  dplyr::mutate(units_sold = round(units_sold, digits = 0))
TidyProdRank <- TidyProdRank %>% 
  bind_cols(UnitsSold)
```

Now we generate the variables for the week-over-week visualizations. 

```{r wow-variables}
KushyWowData <- TidyProdRank %>% 
    dplyr::mutate(
                # day of week
                dow = lubridate::day(invoice_date), 
                # week of the year
                week = lubridate::week(invoice_date),
                # year 
                yr = lubridate::year(invoice_date), 
                # week_year (with floor_date)
                week_year = lubridate::floor_date(invoice_date, 
                                                  unit = "week"),
                # create month
                month = lubridate::month(week_year, 
                                         abbr = TRUE, 
                                         label = TRUE),
                # floor_month
                floor_month = lubridate::floor_date(week_year, 
                                                    unit = "month"))
```

***

Export these data to processed data folder. 

```{r processed-data}
if (!file.exists("data/processed/")) {
  dir.create("data/processed/")
}
# check
fs::dir_tree("data", recurse = FALSE)
```

Time-stamp and export.

```{r export}
# export these with a time stamp
readr::write_csv(base::as.data.frame(KushyWowData), 
                    path = paste0("data/processed/", 
                              base::noquote(lubridate::today()),
                                            "-KushyWowData.csv"))

readr::write_rds(x = KushyWowData, 
                 path = paste0("data/processed/", 
                              base::noquote(lubridate::today()),
                                            "-KushyWowData.rds"))

fs::dir_tree("data/processed/", recurse = FALSE)
```


