---
title: "README - Visualize Cannabis Data"
author: "Martin Frigaard"
date: "current version: `r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
require(tidyverse)
library(plotly)
require(janitor)
require(skimr)
library(mosaic)
library(inspectdf)
library(visdat)
library(DT)
library(hrbrthemes)
# base options ----
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 999999
)
# knitr chunk options ----
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  fig.width = 9, # figure width
  fig.height = 6, # figure height
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/"
) # location of files
# knitr knit settings ----
knitr::opts_knit$set(
  width = 78
)
# set theme
ggplot2::theme_set(theme_ipsum_tw(
  base_family = "DroidSansMono",
  base_size = 10,
  strip_text_size = 9,
  plot_title_family = "TitilliumWeb-SemiBold",
  plot_title_size = 20,
  axis_title_family = "Ubuntu",
  axis_title_size = 13
))
```

## Load the packages

These are the packages we will use to visualize the cannabis data. 

```{r package, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(ggthemes)
library(cluster)
library(factoextra)
library(textshape)
```

## The Online Retail data 

This data came from a UK e-commerce dataset from the [UCI Machine Learning Laboratory](https://archive.ics.uci.edu/ml/datasets/online+retail). 

```{r import-OnlineRetail, message=FALSE, warning=FALSE}
# fs::dir_tree("data")
OnlineRetail <- read_csv("data/raw/Online_Retail.csv") %>% 
    janitor::clean_names(case = "snake")
OnlineRetail %>% dplyr::glimpse(78)
```



```{r wow}
Wow <- OnlineRetail %>% 
    # get the week
    dplyr::select(week_year, quantity) %>%
    dplyr::group_by(week_year) %>%
    dplyr::summarize(week_qty = sum(quantity)) %>%
    dplyr::mutate(prev_week = lag(week_qty,1)) %>%
    dplyr::mutate(wow_quantity = (week_qty - prev_week) / prev_week) %>%
    dplyr::mutate(month = month(week_year, abbr = true, label = true)) %>%
    dplyr::group_by(week_year)
```






ggWoW <- ggplot(data = wow, 
                        mapping = aes(x=week_year, 
                                      y = WoW_Quantity)) +
                                      #color = purchaser_gender)) +
    geom_line() +
    geom_point() +
    theme_fivethirtyeight() +
    theme(axis.title = element_text()) +   
    theme(legend.title = element_blank()) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    facet_wrap(~ month, scales = "free") + 
    ylab('Sales') +
    xlab('Week') +    
    ggtitle('Week Over Week Sales') 

ggWoW 
       

ggQty <- wow %>%
    na.omit() %>%
    ggplot(data=., aes(x = week_year, y = Week_Qty)) + 
    geom_line() +
    geom_smooth() +
    geom_point() +
    #stat_summary(fun.y=mean, geom="bar") + 
    #stat_summary(fun.data=mean_cl_boot, geom="errorbar", width=0.3) +
    theme_fivethirtyeight() +
    theme(axis.title = element_text()) + 
    theme(legend.position = 'none') +
    labs(title = "Quantity Per Week",
         x = "Week", y = "Quantity Sold")


ggQty  


monthly <- ecom %>% 
    select(floor_month, Country, Quantity, UnitPrice) %>%
    mutate(Sales = Quantity*UnitPrice) %>%
    group_by(floor_month,Country) %>%
    summarize(MonthlySales = sum(Sales),
              MedianSale = median(Sales)) %>%
    group_by(floor_month, Country)

ggmonthly <- monthly %>%
    ggplot(data =., 
                mapping = aes(x=floor_month, 
                              y = MonthlySales)) +
    geom_line() +
    geom_point() +
    theme_fivethirtyeight() +
    theme(axis.title = element_text()) +   
    theme(legend.title = element_blank()) +
    scale_y_continuous(labels = scales::dollar_format(accuracy = 1)) +
    facet_wrap(~ Country, scales = "free") + 
    ylab('Sales') +
    xlab('Month') +    
    ggtitle('Monthly Sales') 

ggmonthly 


ggAOV <- monthly %>%
    ggplot(data =., 
           mapping = aes(x=floor_month, 
                         y = MedianSale)) +
    geom_line() +
    geom_point() +
    theme_fivethirtyeight() +
    theme(axis.title = element_text()) +   
    theme(legend.title = element_blank()) +
    scale_y_continuous(labels = scales::dollar_format(accuracy = 1)) +
    facet_wrap(~ Country, scales = "free") + 
    ylab('Sales') +
    xlab('Month') +    
    ggtitle('Median Order Value') 

ggAOV 

customers_sales <- ecom %>%
    select(CustomerID, Quantity, UnitPrice) %>%
    group_by(CustomerID) %>%
    mutate(Sales = (Quantity*UnitPrice)) %>%
    summarize(Sales = sum(Sales),
              MedianSale = median(Sales),
              Quantity = sum(Quantity),
              Median_Quantiy = median(Quantity)) %>%
    group_by(CustomerID)

customers_country <- ecom %>%
    distinct(CustomerID, Country) %>%
    group_by(CustomerID)

customers_sales_country <- customers_sales %>%
    inner_join(customers_country, by = "CustomerID")


customer_sales_clust <- data.frame(customers_sales)
customer_sales_clust <- customer_sales_clust %>%   
                        drop_na() %>%
                        filter(Sales > 0)
summary(customer_sales_clust)
customer_sales_clust <- column_to_rownames(customer_sales_clust) #, var = "CustomerID")

customer_sales_clust <- scale(customer_sales_clust)
k2 <- kmeans(customer_sales_clust, centers =2 , nstart = 25)
k2

fviz_cluster(k2, data = customer_sales_clust)

customer_sales_clust <- data.frame(customer_sales_clust, k2$cluster)



