---
title: "Part 3 - K-Means Cannabis Data"
author: "Martin Frigaard"
date: "current version: `r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
require(knitr)
require(rmdformats)
require(tidyverse)
require(plotly)
require(janitor)
require(skimr)
require(mosaic)
require(inspectdf)
require(visdat)
require(DT)
require(hrbrthemes)
# base options ----
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 999999,
  str = strOptions(
      strict.width = "wrap", 
      vec.len = 3, 
      drop.deparse.attr = TRUE)
)
# knitr chunk options ----
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  fig.width = 9, # figure width
  fig.height = 6.5, # figure height
  warning = FALSE, # show warnings
  message = FALSE,
  size = "small", # size of the text
  fig.path = "figs/"  # location of files
)
# knitr knit settings ----
knitr::opts_knit$set(
  width = 78
)
```

## Load the packages

These are the packages we will use to visualize the cannabis data. 

```{r package, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(ggthemes)
library(cluster)
library(factoextra)
library(textshape)
library(knitr)
library(rmdformats)
library(plotly)
require(janitor)
require(skimr)
library(mosaic)
library(inspectdf)
library(visdat)
library(DT)
library(hrbrthemes)
```

## Import data 

This data came from a UK e-commerce dataset from the [UCI Machine Learning Laboratory](https://archive.ics.uci.edu/ml/datasets/online+retail) and the [kushy cannabis data set](https://github.com/kushyapp/cannabis-dataset). 

```{r import-CannabisWowData, message=FALSE, warning=FALSE}
# fs::dir_tree("data/processed/")
CannabisWowData <- read_csv("data/processed/2020-03-17-CannabisWowData.csv")
```

```{r check-glimpse}
CannabisWowData %>% dplyr::glimpse() 
```

## Create Customer Sales data

These data are the sales by customer id. We create this data frame by grouping on `customer_id`, then create a `sales` variable by multiplying the quantity by the unit price (`quantity * unit_price`).

We also create three variables that aggregate the sales (`sum` and `median`).

```{r CustomerSales}
CustomerSales <- CannabisWowData %>%
    # select the customer_id, quantity, and unit_price
    dplyr::select(id, quantity, unit_price) %>%
    # group it by the customer id
    dplyr::group_by(id) %>%
    # create sales (which is the product of quantity and price)
    dplyr::mutate(sales = (quantity * unit_price)) %>%
    # summarize
    dplyr::summarize(
            sales_sum = sum(sales),
            sales_median = median(sales),
            quantity_sum = sum(quantity),
            quantiy_median = median(quantity)) %>%
    # we group this again by customer id
    dplyr::group_by(id)
CustomerSales %>% dplyr::glimpse(78)
```

## Customer location data

We now join the `CustomerSales` data to a data frame with the distinct customer ids and countries (`CustomersLocation`).

```{r join-CustomersCountry}
CustomersLocation <- CannabisWowData %>%
    dplyr::distinct(id, location) %>%
    dplyr::group_by(id)
# join this to the CustomersLocation
CustomersSalesLocation <- CustomerSales %>%
    dplyr::inner_join(CustomersLocation, 
               by = "id")
# check
CustomersSalesLocation %>% dplyr::glimpse(78)
```

### Create cluster (modeling) data frame

Ok, now we are going to convert the grouped data frame, drop the missing rows, and remove any data where sales were less than zero.

```{r CustomerSalesClust}
# convert to data frame
CustomerSalesClust <- base::data.frame(CustomerSales)

CustomerSalesClust <- CustomerSalesClust %>%  
    # drop na 
                        tidyr::drop_na() %>%
    # remove sales less than 0
                        dplyr::filter(sales_sum > 0)
skimr::skim(CustomerSalesClust)
```

***

## K-Means clustering

Finally, we run a k-means clustering algorithm on the clustered sales data. *What does a k-means clustering do?* Well, here is a great definition on [Medium](https://towardsdatascience.com/understanding-k-means-clustering-in-machine-learning-6a6e67336aa1), 

> the objective of K-means is simple: group similar data points together and discover underlying patterns. To achieve this objective, K-means looks for a fixed number (k) of clusters in a dataset.

The k-means algorithm is an **unsupervised machine learning algorithm.** The steps to perform this are: 

1. Use the `textshape::column_to_rownames()` function on the clustered data frame. This function 

> "*Takes an existing column and uses it as rownames instead. This is useful when turning a data.frame into a matrix. Inspired by the tibble package's `column_to_row` which is now deprecated if done on a `tibble` object. By coercing to a `data.frame` this problem is avoided.*"

We're going to use the non-descript name, `CustomerSalesClust2`. to show what this function is actually doing. The `utils::str()` function gives us some information on what kind of object we've created here.

```{r column_to_rownames}
CustomerSalesClust2 <- textshape::column_to_rownames(CustomerSalesClust) 
utils::str(CustomerSalesClust)
utils::str(CustomerSalesClust2)
```

The `CustomerSalesClust2` took the existing data frame (`CustomerSalesClust`) and removed a column (`customer_id`) and assigned them a rowname `base::rownames()`.

```{r rownames}
base::rownames(CustomerSalesClust2) %>% head()
```


2. The 2nd step is to Scale the new `CustomerSalesClust` with `base::scale()`, which now has been transformed via the `textshape::column_to_rownames()` function. The `scale()` function who's, "*default method centers and/or scales the columns of a numeric matrix.*"

```{r scale}
CustomerSalesClust3 <- base::scale(CustomerSalesClust2)
str(CustomerSalesClust3)
```

Use the `kmeans()` function with the specifications below:

`centers` = A matrix of cluster centres.
`nstart` = 

```{r k2}
k2 <- kmeans(CustomerSalesClust, 
             # the number of cluster centries
             centers = 2 , 
             ## random starts do help here with too many clusters
             ## (and are often recommended anyway!):
             nstart = 25)
k2
```


```{r five-cluster}
fviz_cluster(k2, data = CustomerSalesClust3)
```

```{r CustomerSalesCluser}
CustomerSalesCluser <- base::data.frame(CustomerSalesClust, k2$cluster)
CustomerSalesCluser %>% dplyr::glimpse(78)
```




