---
title: "README - Cannabis Retail Data"
author: "Martin Frigaard"
date: "current version: `r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
require(tidyverse)
library(plotly)
require(janitor)
require(skimr)
library(mosaic)
library(inspectdf)
library(visdat)
library(DT)
# base options ----
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 999999
)
# knitr chunk options ----
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  fig.width = 7, # figure width
  fig.height = 5, # figure height
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/"
) # location of files
# knitr knit settings ----
knitr::opts_knit$set(
  width = 78
)
```


# Cannabis Sales/Marketing Dashboard

The company that I mentioned previously is [Headset](https://www.headset.io/). They recently secured $12.1 M in seed funding which you can read about [here](https://www.newcannabisventures.com/leading-cannabis-analytics-company-headset-closes-12-1-million-series-a/). I agree with you Peter that there is a lot of room in this space and look forward to discovering the most pressing problems we can solve. Other potential competitors in this space to be aware of are:

[Springbig](https://www.springbig.com/) - Read more about them [here](https://www.newcannabisventures.com/springbig-cannabis-dispensary-marketing-engine/).

[DemandLink](https://demandlink.com/solutions/)- Read more about them [here](https://www.newcannabisventures.com/springbig-cannabis-dispensary-marketing-engine/).

I believe Headset is the leader at present. Headset founders also started [Leafly](https://www.leafly.com/), a Yelp-style platform for cannabis.

The landing page for discovery is a great place to start. I have obtained a UK e-commerce dataset from the [UCI Machine Learning Laboratory](https://archive.ics.uci.edu/ml/datasets/online+retail). It can also be found on Kaggle. I've taken the liberty of adding both of you to an RStudio Cloud project I've started with this data. I've cleaned and prepped it for purposes of making some slick time-series graphs, or any cool sample visualizations we can use for the landing page. Feel free to edit an collaborate as you see fit.

## Project folder structure 

```{r tree}
fs::dir_tree(".")
```

### Data visualizations 

The data visualizations are built using the ggplot2 and plotly package. 


